---
title: "A Quick Introduction to R Markdown"
author: "Paul Staab"
date: "09.10.2014"
output: html_document
---

__R Markdown__ is a simple markup language that can be used to create reports about an 
stastistical analysis directly _while executing the analysis_. It is a quite convenient way to record
the analysis -- especially in case you need to explain what you did in, say, three years from 
now -- and can also be used to convince other researchers (your supervisor, reviewers,...) 
that your results are solid. It is also designed to ensure a high level of reproducibility of
your analysis, which is always a major concern when conducting research.


Creating an R Markdown Document
-------------------------------
In `RStudio`, such reports can be conveniently  created by using the 

    File -> New File -> R Markdown
    
menu or by selecting `R Markdown` after clicking the `New File Button` 
in the toolbar. 


Using Markdown
--------------
Markdown is a very simplistic format for writing text. RStudio has a `Quick Reference`
build-in that explains how you add Headers, Lists, Images, Links etc to your document.
It available by clicking of the `?` button at the top of the document.


Adding R Commands
-----------------
The main think about R Markdown is that you can add chunks of R code to the document.
This code get automatically executed when you build ('knit') the document, and all 
of it's output is inserted into the document, including plots.

Code chunks always start with the line

> ```{r}

where addition options can be inserted after the `r`. Chunks end with

> ```

and everything in-between is executed with R. For instance a chunk
can load the `iris` data set, print a summary and produce a plot:

```{r}
data(iris)
head(iris, n=4)
library(ggplot2)
ggplot(iris, aes(Sepal.Length, Sepal.Width, color=Species)) + geom_point()
```

When knitting a document, first a fresh environment is created to facilitate 
reproducibility of the report. That in particular implies that you must reload
all packages you use with the `library()` command before you can use a function
of the package. Afterwards all chunks are executed sequentially in this
environment, so the next chunk can use objects created before:

```{r}
x <- 1:10
```

And in the next chunk, `x` is still available:

```{r}
sum(x)
```

R Markdown documents can be knitted to html, pdf (requires LaTeX installed) and Word documents.
Use the `Knit` button in the toolbar above the document.

Note that you can also send R commands to the console (`Run` or `CRTL - ENTER`) when writing
an R Markdown file. As long as the cursor is inside an R chunk, `RStudio` behaves as it would
when working with an R script.



Tips
-------------------
A common recommendation is that if you conduct an important analysis for a paper, you actually
**two** reports. One for yourself that contains and explains everything that you tried, and a
second polish one that just contains the things that worked and are important for the paper. The 
first one is for yourself to look things up later, the latter to convince the reviewers. 


Troubleshooting
-------------------
- If you run into problems, try understanding the error message you get, and in particular look if it 
contains a line number in which the error occurs. Double check that line, and the previous one. 
- Using a recent version of `RStudio` and `R` is usually a good idea.
- Try googling the error message.
- If nothing helps, you often get high quality answers on sites like 
  [Stack Overflow](https://stackoverflow.com/questions/tagged/r). 
  Spend some time exactly describing your problem as detailed as possible.